---
title: "The Game is Afoot & Poems of Mihai Eminescu"
author: "Jessica Randall"
date: "9/4/2020"
output: html_document
---

This notebook follow's Julia Silge's topic modeling tutorial [video](https://youtu.be/evTuL-RcRpc)
and [blog post](https://juliasilge.com/blog/sherlock-holmes-stm/) and expands on 
it with an example of topic modeling using Mihai Eminescu's collection "Poems" in 
the original Romanian. 

Additional Resources:
[Tidy Text Mining](https://www.tidytextmining.com/tidytext.html)
[Supervised Machine Learning for Text Analysis in R](https://smltar.com)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(
  "tidymodels", "tidyverse", "gutenbergr",
  "tidytext", "stringr", "stm", "quanteda", "reshape2",
  "furrr", "ggthemes", "kableExtra", "styler"
)

sherlock_raw <- gutenberg_download(1661)
```

### Data prep {-} 

```{r prep, message=FALSE, warning=FALSE}

sherlock <- sherlock_raw %>%
  mutate(story = ifelse(str_detect(text, "ADVENTURE"),
    text,
    NA
  )) %>%
  fill(story) %>%
  filter(story != "THE ADVENTURES OF SHERLOCK HOLMES") %>%
  mutate(story = factor(story, levels = unique(story)))

tidy_sherlock <- sherlock %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  filter(word != "holmes")

tidy_sherlock %>%
  count(word, sort = TRUE)
```

### Explore tf-idf {-}

Which words are most important in each story?

tf-idf: "identifies words that are important to a document in a collection of 
documents; in this case."The statistic tf-idf is intended to measure how 
important a word is to a document in a collection (or corpus) of documents, 
for example, to one novel in a collection of novels or to one website in a 
collection of websites.

decreases the weight for commonly used words and increases the weight for words 
that are not used very much in a collection of documents. This can be combined 
with term frequency to calculate a term’s tf-idf (the two quantities 
multiplied together), the frequency of a term adjusted for how rarely it is used."

see : https://www.tidytextmining.com/tfidf.html

```{r tf_idf, fig.height=7, fig.width=7}

tidy_sherlock %>%
  count(story, word, sort = TRUE) %>%
  bind_tf_idf(word, story, n) %>%
  group_by(story) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = story)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~story, scales = "free") +
  coord_flip()
```

### Topic modeling implementation {-}

Structural topic model, em algorithm, k will vary by document, 
slightly computationally intensive

on k :The most important user input in parametric topic models is the number of 
topics. There is no right answer to the appropriate number of topics. More 
topics will give more fine-grained representations of the data at the potential 
cost of being less precisely estimated." - documentation for stm


```{r quanteda, message=FALSE, echo=FALSE, warning=FALSE}

sherlock_dfm <- tidy_sherlock %>%
  count(story, word, sort = TRUE) %>%
  cast_dfm(story, word, n)

topic_mod <- stm(sherlock_dfm, K = 6, seed = 22310)
```

### Plot the beta and gamma matrices {-}

-Which words contribute the most to each topic?
-How much does each topic contribute to a given document?
-How likely is this document to belong to this topic?

```{r plot, fig.height=7, fig.width=7}

td_beta <- tidy(topic_mod)

td_beta %>%
  group_by(topic) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = topic)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  theme_few() +
  scale_color_colorblind()

td_gamma <- tidy(topic_mod,
  matrix = "gamma",
  document_names = rownames(sherlock_dfm)
)


td_gamma %>%
  ggplot(aes(gamma, fill = as.factor(topic))) +
  geom_histogram(show.legend = TRUE) +
  facet_wrap(~topic, ncol = 3) +
  theme_gdocs() +
  scale_fill_colorblind()
```

### Trying on my own with Mihai Eminescu's "Poems" {-}

```{r load_jr}

poezii_raw <- gutenberg_download(35323)
```

### Data prep {-} 

- Only taking first half of poems of the raw corpus for better visualization
- Removed editors note, table of contents, 
- Used poem titles
- Pulled out Romanian stopwords, the word "poems", notes, numbers

```{r prep_jr, warning=FALSE, message=FALSE}

poezii <- poezii_raw[151:12887, ] %>%
  mutate(poem = ifelse(str_detect(text, "[:upper:]+$"), text, NA)) %>%
  fill(poem) %>%
  mutate(poem = factor(poem, levels = unique(poem)))

stop_snow <- as.data.frame(stopwords("ro", source = "snowball")) %>%
  rename(word = `stopwords("ro", source = "snowball")`)
stop_nltk <- as.data.frame(stopwords("ro", source = "nltk")) %>%
  rename(word = `stopwords("ro", source = "nltk")`)
stop_iso <- as.data.frame(stopwords("ro", source = "stopwords-iso")) %>%
  rename(word = `stopwords("ro", source = "stopwords-iso")`)
numbers <- c(
  "I", "	II", "III", "IV", "V", "VI",
  "VII", "VIII", "11", "1652", "38", "5", "6", "9"
)

stopwords_ro <- left_join(stop_iso, stop_nltk, by = "word") %>%
  left_join(stop_snow, by = "word")

tidy_poezii <- poezii %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text) %>%
  anti_join(stopwords_ro) %>%
  filter(word != "poezii") %>%
  filter(!word %in% numbers) %>%
  filter(!poem %in% numbers) %>%
  filter(poem != "NOTE")

tidy_poezii %>%
  count(word, sort = TRUE)

```

"Eyes", World", "Sweet", and "Life" top the list of most frequently occuring words,
sounds like poetry!

### Explore tf-idf {-}

Which words are most important?

```{r tf_idf_jr, fig.height=10, fig.width=10, message=FALSE}

poezii_tf_idf <- tidy_poezii %>%
  count(poem, word, sort = TRUE) %>%
  arrange(poem, word, n) %>%
  slice(1:2583) %>%
  bind_tf_idf(word, poem, n) %>%
  group_by(poem) %>%
  top_n(5) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf))

poezii_tf_idf %>%
  ggplot(aes(word, tf_idf, fill = poem)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 3) +
  coord_flip() +
  theme(strip.text = element_text(size = 11)) +
  labs(
    x = NULL, y = "tf-idf",
    title = "Highest tf-idf words in 15 of Mihai Eminescu's Poems",
    subtitle = "Individual poems focus on different themes and visual metaphors"
  ) +
  theme_few()
```

"The Grave of Aron Pumunul" lists the 5 most important words as "whispers", 
"whispering", "extinguished", "follow", "field". 

"A Ride at Dawn" lists the 5 most important words as "kiss", "gentle", "thin", 
"cleave", and "wash". 

"From Abroad" lists the 5 most important words as "date", "whispered", "native", 
"groan", and "clouds".

"In Bucovina" lists the 5 most important words as "Bucovina", "step", "hum", 
"tremble", and "pass".

"The Hope" lists the 5 most important words as "hope" "I want", "love", 
"dearest", and "you want".

"The Mysteries of the Night" lists the 5 most important words as "many", 
"gentle"(diminutive), "sighing", "Cantu", and "two". 

"What I Wish From You, Beloved Romania" lists the 5 most important words as 
"Romania" "I wish", "for you", "glory", and "strength". 

"The Heliad" lists the 5 most important words as "garland", "sylphs", "heliads", 
"cloud", and "curls".

"The Artist" lists the 5 most important words as "take", part of the word
"coming true", "embodied", "flower", and "deify".

"The Love of a Marble" lists the 5 most important words as "I love" "your", 
"ocean", "love", and "tell".

"Junii Corupti"* lists the 5 most important words as "forget", "dry", "death" 
or "dying", "arise", and "strength". 

"At the Death of Prince Stirbey" lists the 5 most important words as "move", 
"Stirbey", "light", "to see", and "hearts". 

"Venus and Madonna" lists the 5 most important words as "women", "lost",
"Madonna", "saw", and "demon". 

"The Epigones" lists the 5 most important words as "gold", "holy", "all", 
"sense", and "death". 

"Guardian Angel" lists the 5 most important words as "across", "you are", 
"eyelashes", "keep watch", and "guard". 

*the best translation of this title I can think of would be something like 
"The Corrupted" but this loses the fact that he's speaking specifically about 
young people, suggestions appreciated!

### Topic modeling implementation {-}

*set up our dfm*

```{r quanteda_jr, message=FALSE, echo=FALSE, warning=FALSE}

poezii_dfm <- tidy_poezii %>%
  count(poem, word, sort = TRUE) %>%
  cast_dfm(poem, word, n)
```

*evaluate which k is the best fit for this corpus*

Julia knew that a k of 6 was a good choice for her data but since I don't know 
what would be the best fit for this data, let's find out.

This section adapted from Julia's post ["Training, evaluating, and interpreting 
topic models"](https://juliasilge.com/blog/evaluating-stm/)

```{r choosing_k, message=FALSE, echo=FALSE}

plan(multiprocess)

multi_mods <- tibble(K = c(12, 14, 15, 16, 17, 18)) %>%
  mutate(topic_model = future_map(K, ~ stm(poezii_dfm,
    K = .,
    verbose = FALSE,
    seed = 22310
  )))

heldout <- make.heldout(poezii_dfm)

k_result <- multi_mods %>%
  mutate(
    exclusivity = map(topic_model, exclusivity),
    semantic_coherence = map(topic_model, semanticCoherence, poezii_dfm),
    eval_heldout = map(topic_model, eval.heldout, heldout$missing),
    residual = map(topic_model, checkResiduals, poezii_dfm),
    bound = map_dbl(topic_model, function(x) max(x$convergence$bound)),
    lfact = map_dbl(topic_model, function(x) lfactorial(x$settings$dim$K)),
    lbound = bound + lfact,
    iterations = map_dbl(topic_model, function(x) length(x$convergence$bound))
  )
```

### Visualize fit metrics {-}

```{r viz_assessment, echo=FALSE, message=FALSE, warning=FALSE}

k_result %>%
  transmute(K,
    `Lower bound` = lbound,
    Residuals = map_dbl(residual, "dispersion"),
    `Semantic coherence` = map_dbl(semantic_coherence, mean),
    `Held-out likelihood` = map_dbl(eval_heldout, "expected.heldout")
  ) %>%
  gather(Metric, Value, -K) %>%
  ggplot(aes(K, Value, color = Metric)) +
  geom_line(size = 1.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(
    x = "K (number of topics)",
    y = NULL,
    title = "Model diagnostics by number of topics",
    subtitle = "These diagnostics indicate that a good number of topics 
       would be around X"
  ) +
  theme_gdocs() +
  scale_color_colorblind()

k_result %>%
  select(K, exclusivity, semantic_coherence) %>%
  filter(K %in% c(12, 14, 15, 16, 17, 18)) %>%
  unnest(cols = c(exclusivity, semantic_coherence)) %>%
  mutate(K = as.factor(K)) %>%
  ggplot(aes(semantic_coherence, exclusivity, color = K)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(
    x = "Semantic coherence",
    y = "Exclusivity",
    title = "Comparing exclusivity and semantic coherence",
    subtitle = "Models with fewer topics have higher semantic coherence 
       for more topics, but lower exclusivity"
  ) +
  theme_gdocs() +
  scale_color_colorblind()
```

The held-out likelihood is the highest at 14 topics and the residuals are
lowest around 14. Coherence also peaks at 14 so it seems like 14 topics would be
the best fit for this data. Looking more closely at coherence we see that the 
ks of 14 have the highest exclusivity and highest coherence. I'll choose 14 for 
this analysis.

### Plot results {-}

*Beta*

```{r plotbeta, fig.height=12, fig.width=12}

poezii_mod <- k_result %>%
  filter(K == 14) %>%
  pull(topic_model) %>%
  .[[1]]

td_beta_p <- tidy(poezii_mod)

td_beta_p %>%
  group_by(topic) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = topic)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(
    x = NULL, y = expression(beta),
    title = "Highest word probabilities for each topic",
    subtitle = "Different words are associated with different topics"
  ) +
  theme_few() +
  scale_color_colorblind()
```
Many of the topics have overlapping words like "sweet", "stars", "gold", "moon",
"eyes", "world", "your", "life", "light" etc... again, sounds like poetry!

*Gamma*

```{r plotgamma, fig.height=7, fig.width=7, warning=FALSE, message=FALSE}

td_gamma_p <- tidy(poezii_mod,
  matrix = "gamma",
  document_names = rownames(poezii_dfm)
)

td_gamma_p %>%
  ggplot(aes(gamma, fill = as.factor(topic))) +
  geom_histogram(show.legend = TRUE) +
  facet_wrap(~topic, ncol = 3) +
  labs(
    title = "Distribution of document probabilities for each topic",
    subtitle = "Each topic is associated with 1-22 stories",
    y = "Number of poems", x = expression(gamma)
  ) +
  theme_gdocs() +
  scale_color_colorblind()

top_terms <- td_beta_p %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(7, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term), .groups = "keep") %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>%
  unnest(cols = c(topic, terms))

gamma_terms <- td_gamma_p %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma), .groups = "keep") %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms, by = "topic") %>%
  mutate(
    topic = paste0("Topic ", topic),
    topic = reorder(topic, gamma)
  )

gamma_terms %>%
  top_n(7, gamma) %>%
  ggplot(aes(topic, gamma, label = terms, fill = topic)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3) +
  coord_flip() +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 0.2),
    labels = percent_format()
  ) +
  labs(
    x = NULL, y = expression(gamma),
    title = "7 topics by prevalence in the Poems corpus",
    subtitle = "With the top words that contribute to each topic"
  ) +
  theme_few() +
  scale_color_colorblind() +
  theme(
    plot.title = element_text(size = 14),
    plot.subtitle = element_text(size = 12))

gamma_terms %>%
  select(topic, gamma, terms) %>%
  kable(
    digits = 3,
    col.names = c("Topic", "Expected topic proportion", "Top 7 terms")
  )

```


Topics 6 and 13 share the words "eyes" and "sweet" but 6 focuses more on 
the subject of the poem and uses words like "you are", "your", and "mouth" while
13 has more physical words like "hand", "heart/soul", "life" but also "light, 
"shadow", and "country".

Topic 8 includes words related to night time like "moon", "stars", and "silver".

Many of the topics seem to have grouped words which end in similar ways.
Romanian uses word endings to denote concepts like ownership "my", "your", 
"ours", plurality similar to the English ending "-s", or articles like "a" or 
"the" so in this case stemming may be a good choice. Might want to come back to
this and try stemming but I'm conflicted since I know it can often hurt.

I'm not sure how much this type of modeling told me about the text but I imagine
if I had a huge corpus of text that covered a wide variety of topics then this 
might be more informative. It definitely confirms my impression of Eminescu
as being a very sentimental, romantic poet. Maybe a sentiment analysis would be
a good idea for his works?

This was a really fun exercise in topic modeling and I'm grateful to Julia Silge
for her super informative blog!