---
title: "The Game is Afoot & Poems of Mihai Eminescu"
author: "Jessica Randall"
date: "9/4/2020"
output: html_document
---

This notebook follow's Julia Silge's topic modeling tutorial [video](https://youtu.be/evTuL-RcRpc)
and [blog post](https://juliasilge.com/blog/sherlock-holmes-stm/) and expands on 
it with an example of topic modeling using Mihai Eminescu's collection "Poems" in 
the original Romanian. 

Additional Resources:
[Tidy Text Mining](https://www.tidytextmining.com/tidytext.html)
[Supervised Machine Learning for Text Analysis in R](https://smltar.com)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load("tidymodels", "tidyverse","gutenbergr", 
               "tidytext", "stringr", "stm", "quanteda", "reshape2")

sherlock_raw <- gutenberg_download(1661)
```

## data prep {-} 

```{r prep, message=FALSE, echo=FALSE}

sherlock <- sherlock_raw %>%
  mutate(story = ifelse(str_detect(text, "ADVENTURE"), 
                        text, 
                        NA)) %>%
  fill(story) %>%
  filter(story != "THE ADVENTURES OF SHERLOCK HOLMES") %>%
  mutate(story = factor(story, levels = unique(story)))

tidy_sherlock <- sherlock %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  filter(word != "holmes")

tidy_sherlock %>%
  count(word, sort = TRUE)

```

## explore tf-idf {-}

Which words are most important in each story?

tf-idf: "identifies words that are important to a document in a collection of 
documents; in this case."The statistic tf-idf is intended to measure how 
important a word is to a document in a collection (or corpus) of documents, 
for example, to one novel in a collection of novels or to one website in a 
collection of websites.

decreases the weight for commonly used words and increases the weight for words 
that are not used very much in a collection of documents. This can be combined 
with term frequency to calculate a term’s tf-idf (the two quantities 
multiplied together), the frequency of a term adjusted for how rarely it is used."

see : https://www.tidytextmining.com/tfidf.html

```{r tf_idf, include=FALSE}

tidy_sherlock %>%
  count(story, word, sort = TRUE) %>%
  bind_tf_idf(word, story, n) %>%
  group_by(story) %>%
  top_n(10) %>%
  ungroup %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = story)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~story, scales = "free") +
  coord_flip()

```

## topic modeling implementation

structural topic model, em algorithm, k will vary by document, 
slightly computationally intensive

on k :The most important user input in parametric topic models is the number of 
topics. There is no right answer to the appropriate number of topics. More 
topics will give more fine-grained representations of the data at the potential 
cost of being less precisely estimated." - documentation for stm


```{r quanteda, include=FALSE}

# create dfm
sherlock_dfm <- tidy_sherlock %>%
  count(story, word, sort = TRUE) %>%
  cast_dfm(story, word, n)

# fit the model
topic_mod <- stm(sherlock_dfm, K = 6, seed = 22310)

summary(topic_mod)

```

## plot the beta matrix

beta probs: Which words contribute the most to each topic?

gamma probs: How much does each topic contribute to a given document?
How likely is this document to belong to this topic?
this is a really good example of a model that did a good job putting stories
into topics topics are all very strongly associated with one topic, not always 
common, this is likely because this data set is very small

```{r plot include=FALSE}

td_beta <- tidy(topic_mod)

td_beta %>%
  group_by(topic) %>%
  top_n(10) %>%
  ungroup %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = topic)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~topic, scales = "free") +
  coord_flip()

td_gamma <- tidy(topic_mod, matrix = "gamma",
             document_names = rownames(sherlock_dfm))


td_gamma %>%
  ggplot(aes(gamma, fill = as.factor(topic))) +
  geom_histogram(show.legend = TRUE) +
  facet_wrap(~topic, ncol = 3)

```

## trying on my own with Mihai Eminescu's "Poems"

```{r load_jr}

poezii_raw <- gutenberg_download(35323)

```

## data prep {-} 

only taking first half of poems of the raw corpus for better visualization
removed editors note, table of contents, 
used poem titles,
pulled out romanian stopwords, the word "poems", notes, numbers

```{r prep_jr, message=FALSE, echo=FALSE}

poezii <- poezii_raw[151:12887,] %>%
  mutate(poem = ifelse(str_detect(text, "[:upper:]+$"), text, NA)) %>%
  fill(poem) %>%
  mutate(poem = factor(poem, levels = unique(poem)))

stop_snow <- as.data.frame(stopwords("ro", source = "snowball")) %>%
  rename(word = `stopwords("ro", source = "snowball")`)
stop_nltk <- as.data.frame(stopwords("ro", source = "nltk")) %>%
  rename(word = `stopwords("ro", source = "nltk")`)
stop_iso <- as.data.frame(stopwords("ro", source = "stopwords-iso")) %>%
  rename(word = `stopwords("ro", source = "stopwords-iso")`)
numbers <- c("I", "	II", "III", "IV", "V", "VI", 
             "VII", "VIII", "11", "1652","38", "5", "6", "9")

stopwords_ro <- left_join(stop_iso, stop_nltk, by = "word") %>%
  left_join(stop_snow, by = "word")

tidy_poezii <- poezii %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text) %>%
  anti_join(stopwords_ro) %>%
  filter(word != "poezii") %>%
  filter(!word %in% numbers) %>%
  filter(!poem %in% numbers) %>%
  filter(poem != "NOTE")

tidy_poezii %>%
  count(word, sort = TRUE)

```

## explore tf-idf {-}

Which words are most important?

filtered words with tf_idf > 0.025 so that all words would show up 
clearly on the graph

```{r tf_idf_jr, include=FALSE}

#include only first 15 poems in sample

poezii_tf_idf <- tidy_poezii %>%
  count(poem, word, sort = TRUE) %>%
  arrange(poem, word, n) %>%
  slice(1:2583) %>%
  bind_tf_idf(word, poem, n) %>%
  group_by(poem) %>%
  top_n(5) %>%
  ungroup %>%
  mutate(word = reorder(word, tf_idf))

poezii_tf_idf %>%
  ggplot(aes(word, tf_idf, fill = poem)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~poem, scales = "free", ncol = 3) +
  coord_flip() +
  theme(strip.text = element_text(size = 11)) +
  labs(x = NULL, y = "tf-idf",
         title = "Highest tf-idf words in 15 of Mihai Eminescu's Poems",
         subtitle = "Individual poems focus on different themes and visual metaphors")

```

"The Grave of Aron Pumunul" lists the 5 most important words as "whispers", 
"whispering", "extinguished", "follow", "field". 

"A Ride at Dawn" lists the 5 most important words as "kiss", "gentle", "thin", 
"cleave", and "wash". 

"From Abroad" lists the 5 most important words as "date", "whispered", "native", 
"groan", and "clouds".

"In Bucovina" lists the 5 most important words as "Bucovina", "step", "hum", 
"tremble", and "pass".

"The Hope" lists the 5 most important words as "hope" "I want", "love", 
"dearest", and "you want".

"The Mysteries of the Night" lists the 5 most important words as "many", 
"gentle"(diminutive), "sighing", "Cantu", and "two". 

"What I Wish From You, Beloved Romania" lists the 5 most important words as 
"Romania" "I wish", "for you", "glory", and "strength". 

"The Heliad" lists the 5 most important words as "garland", "sylphs", "heliads", 
"cloud", and "curls".

"The Artist" lists the 5 most important words as "take", part of the word
"coming true", "embodied", "flower", and "deify".

"The Love of a Marble" lists the 5 most important words as "I love" "your", 
"ocean", "love", and "tell".

"Junii Corupti"* lists the 5 most important words as "forget", "dry", "death" 
or "dying", "arise", and "strength". 

"At the Death of Prince Stirbey" lists the 5 most important words as "move", 
"Stirbey", "light", "to see", and "hearts". 

"Venus and Madonna" lists the 5 most important words as "women", "lost",
"Madonna", "saw", and "demon". 

"The Epigones" lists the 5 most important words as "gold", "holy", "all", 
"sense", and "death". 

"Guardian Angel" lists the 5 most important words as "across", "you are", 
"eyelashes", "keep watch", and "guard". 


*the best translation of this title I can think of would be something like 
"The Corrupted" but this loses the fact that he's speaking specifically about 
young people, suggestions appreciated!

## topic modeling implementation

structural topic model, em algorithm, k will vary by document, 
slightly computationally intensive

on k :The most important user input in parametric topic models is the number of 
topics. There is no right answer to the appropriate number of topics. More 
topics will give more fine-grained representations of the data at the potential 
cost of being less precisely estimated." - documentation for stm


```{r quanteda, include=FALSE}

# create dfm
poezii_dfm <- tidy_poezii %>%
  count(poem, word, sort = TRUE) %>%
  cast_dfm(poem, word, n)

# fit the model
poezii_mod <- stm(poezii_dfm, K = 18, seed = 22310)

summary(poezii_mod)
```

## plot the beta matrix

beta probs: Which words contribute the most to each topic?

gamma probs: How much does each topic contribute to a given document?
How likely is this document to belong to this topic?
this is a really good example of a model that did a good job putting stories
into topics topics are all very strongly associated with one topic, not always 
common, this is likely because this data set is very small

```{r plot include=FALSE}

td_beta_p <- tidy(poezii_mod)

td_beta_p %>%
  group_by(topic) %>%
  top_n(10) %>%
  ungroup %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = topic)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics")

td_gamma_p <- tidy(poezii_mod, matrix = "gamma",
             document_names = rownames(poezii_dfm))

td_gamma_p %>%
  ggplot(aes(gamma, fill = as.factor(topic))) +
  geom_histogram(show.legend = TRUE) +
  facet_wrap(~topic, ncol = 3) +
  labs(title = "Distribution of document probabilities for each topic",
       subtitle = "Each topic is associated with 1-3 stories",
       y = "Number of stories", x = expression(gamma))

```

The graph of beta probabilities suggests that there are 9 topics within 12 poems.
This seems reasonable since we might expect that a poet would write at least a
few poems with somewhat overlapping topics. 

The first topic includes words largely from "The Corrupted" and this seems
reasonable as the mood of this poem seems much darker than the rest of those 
included in this sample. The second topic is largely comprised of words from 
"What I Wish From You, Beloved Romania" and this was the only poem that seemed
to be about Romania specifically. Interestingly this also includes the word
"sweet" which we'll explore in other topics.

The third topic is largely from "The Artist", the fourth from "Mysteries of the 
Night", and the fifth topic is where we start to see overlapping words from a 
few poems and in fact have three topics which rate the word "sweet" as most 
important. 

Each of the topics which rate "sweet" as most important seems to suggest a 
different type of sweetness. The fifth topic rates words like "sweet", "gentle", 
"forehead", "song", and "curls" as most important so it seems to be picking up 
romantic sweetness. The sixth topic rates "sweet", "hope", "wish", "gold", 
"forget" as most important so it seems to be picking up on the sense of the 
sweetness of fulfilled dreams or wishes and upon closer examination many of these
words are present in "The Hope". The ninth topic rates words like 
"sweet", "love", "crying", "extinguished", "whispering", "whispers" so 
it's picking up the bittersweetness likely present in many Eminescu poems.
It likely comes as no surprise to those who are even passingly familiar with
Eminescu's poetry that many of the topics are quite sentimental but this is a ver



















